Assessing Phonon Properties of Shape Memory Alloys using Graph Neural Network Potentials

### 2024-01-04 src/tex/ms.tex
Changes to Calculators and Conda Env. Additions.

\subsection{Workflow}
\label{sec:workflow}
The ground-state structures for the NiTi phases \textit{B2}, \textit{B19}, \textit{B19'}, and \textit{BCO} are generated using the atomic simulation environment (ASE) package \cite{Larsen2017}. This requires that the unit cell parameters, basis atoms, and spacegroup number be provided. The initial unit cell parameters and basis corrdinates for these structures have been taken from the materials project \cite{Jain2013} and several references \cite{Haskins2016,Kadkhodaei2018}. The initial structures for NiTi are shown in fig~\ref{fig:niti_structures} and are adapted for PtTi as well given same spacegroup symmetries. The energy and forces on the atoms are determined using the ASE calculator interfaces for EAM, MEAM, M3GNet, CHGNet, and MACE.\cite{Mutter2010,Zhong2011,Ko2015,Kim2017,Chen2022,Deng2023,Batatia2022} The EAM and MEAM potential interfaces are achieved through the use of LAMMPS python library \cite{Thompson2022} and ASE. The ALIGNN ASE calculator interface was also implemented, however, we found that the predictions for NiTi and PtTi to be unreliable are not included in this study.\par

@@

Prior to any property calculations or dynamics the structures are optimized using the FIRE\cite{Bitzek2006} local optimizer with the forces converged to $2\cdot 10^{-3}$eV/\AA. We observed that other local optimization routines find the same ground-state structures. To ensure that the unit cell parameters were adjusted along with the atomic coordinates and that the crystal symmetry is not broken, we utlize a constraint and filter during the optimization process. \par

@@

\subsubsection{Phonon Bandstructures}
Phonon calculations are performend using finite-difference method via small displacement of atoms in a reference cell.\cite{Alfe2009} In the small displacement method, atoms in the primitive cell are slightly displaced to calculate forces on other atoms, constructing the force constant matrix, which, when used with the dynamical matrix, allows computing phonon frequencies at any q-vector in the Brillouin zone. This requires typically requires the use of a supercell of the original unit cell. In order to ensure convergence of this method, one needs to sweep different supercell sizes and atomic displacements. This process is done for each interatomic potential and structure. The supercells used in this work  have a repeat in any one direction ranging from $5-8$ and depends on the crystal system. The atomic displacement ranges from $0.03-0.26$\AA\,and was found to strongly depend on interatomic potential and crystal system. Details on the settings are provided in the appendix~\ref{sec:phonon_calc_config}. \par

@@

@@

@@

\subsection{NiTi: Ground State Structures}
\label{subsec:nitigs}

%\comment{Display results in tables}

@@

\subsection{PtTi: Equation of States}

@@

\section*{Reproducibilty and Data Availability}
This study was carried out using the reproducibility software \href{https://github.com/showyourwork/showyourwork}{\showyourwork}, which leverages continuous integration to programmatically download the data from \href{https://zenodo.org/}{zenodo.org}, create the figures, and compile the manuscript. Each figure caption contains two links: one to the dataset stored on zenodo used in the corresponding figure, and the other to the script used to make the figure (at the commit corresponding to the current build of the manuscript). The git repository associated to this study is publicly available at \url{https://github.com/stefanbringuier/SMA_Phonons_GNNIP} and allows anyone to re-build the entire manuscript. The datasets generated by this paper are stored at \url{xyz}.

@@

\appendix
\subsection{Phonon Calculation Configurations}
\label{sec:phonon_calc_config}

>>>>

\subsection{Workflow}
\label{sec:workflow}
To promote reproducible research and align with findable, accessible, interoper-able and reusable (FAIR) principles\cite{Walsh2024}, we utilize \showyourwork python workflow\cite{Luger2021} and several open source python packages. Instructions for reproducing calculation data, figures, tables, and manuscript are provided in appendix~\ref{sec:appx_reproduce} and corresponding github repo\footnote{\url{https://github.com/stefanbringuier/SMA_Phonons_GNNIP}}. \par

\subsection{Calculation Details}
\label{sec:calc_details}
The ground-state structures for the NiTi phases \textit{B2}, \textit{B19}, \textit{B19'}, and \textit{BCO} are generated using the atomic simulation environment (ASE) package \cite{Larsen2017}. This requires that the unit cell parameters, basis atoms, and spacegroup number be provided. The initial unit cell parameters and basis corrdinates for these structures have been taken from the materials project \cite{Jain2013} and several references \cite{Haskins2016,Kadkhodaei2018}. The initial structures for NiTi are shown in fig~\ref{fig:niti_structures} and are adapted for PtTi as well given same spacegroup symmetries. The energy and forces on the atoms are determined using the ASE calculator interfaces for EAM, MEAM, M3GNet, CHGNet, and MACE.\cite{Mutter2010,Zhong2011,Ko2015,Kim2017,Chen2022,Deng2023,Batatia2022} The EAM and MEAM potential interfaces are achieved through the use of LAMMPS python library \cite{Thompson2022} and ASE. The ALIGNN ASE calculator interface was also implemented, however, we found that the predictions for NiTi and PtTi to be unreliable are not included in this study.\par

@@

Prior to any property calculations or dynamics the structures are optimized using the FIRE\cite{Bitzek2006} local optimizer with the forces converged to $2\cdot 10^{-3}$eV/\AA. We observed that other local optimization routines find the same ground-state structures. To ensure that the unit cell parameters were adjusted along with the atomic coordinates and that the crystal symmetry is not broken, we utlize a constraint and filter during the optimization process. The optimized lattice parameters as well as basis atoms for all structures and models are provided in appendix \ref{sec:opt_structures}. \par

@@

\subsubsection{Phonon Bandstructures}
Phonon calculations are performend using finite-difference method via small displacement (i.e., harmonic approximation) of atoms in a reference cell \cite{Alfe2009}. In the small displacement method, atoms in the primitive cell are slightly displaced to calculate forces on other atoms, constructing the force constant matrix, which, when used with the dynamical matrix, allows computing phonon frequencies at any q-vector in the Brillouin zone. This requires typically requires the use of a supercell of the original unit cell. In order to ensure convergence of this method, one needs to sweep different supercell sizes and atomic displacements. This process is done for each interatomic potential and structure. The supercells used in this work  have a repeat in any one direction ranging from $5-8$ and depends on the crystal system. The atomic displacement ranges from $0.03-0.26$\AA\,and was found to strongly depend on interatomic potential and crystal system. Details on the settings are provided in the appendix~\ref{sec:phonon_calc_config}. \par

@@

Similar to the phonon calculations the elastic constants within the harmonic approximation the elastic constants can be determined from linear-response of strain-stress calculations. Therefore, the calculation requires the use of crystal symmetry along with atomic displacements to determine respective elastic constant componenets. We use the \textit{Elastic}\cite{Jochym2018} python package which provides routines for generating the proper atomic displacement matrix given specific crystal spacegroup. The elastic constants are determined by solving a set of linear equations using singular value decomposition and given calculated strains and stress. The displacements occur with a min and max of 0.5\% and a total of 10 points are used. \par

@@

The M3GNet has been implemented in the MatGL framework and we utlize the \texttt{M3GNet-MP-2021.2.8-PES} pretrained model.

For CHGNet, we have used the \texttt{v0.3} series pretrained model.

%The MACE potential pretrained \texttt{2023-12-03-mace-128-L1_epoch-199.model} checkpoint file.

@@

@@

\variable{output/Table_NiTi_Elastic_Constants.tex}

\subsection{PtTi: Equation of States}
\label{sec:ptti_eos}

@@

\section*{Reproducibilty and Data Availability}
This study was carried out using the reproducibility software \href{https://github.com/showyourwork/showyourwork}{\showyourwork}\cite{Luger2021}, which leverages continuous integration to programmatically download the data from \href{https://zenodo.org/}{zenodo.org}(if needed), create the figures, and compile the manuscript. Each figure caption contains two links: one to the dataset stored on zenodo used in the corresponding figure, and the other to the script used to make the figure (at the commit corresponding to the current build of the manuscript). The git repository associated to this study is publicly available at \url{https://github.com/stefanbringuier/SMA_Phonons_GNNIP} and allows anyone to re-build the entire manuscript.\par

@@

\newpage
\appendix

\subsection{Workflow Reproducibility}
\label{sec:appx_reproduce}

TODO

\subsection{NiTi: Ground State Structures}
\label{sec:opt_structures}

\subsection{Phonon Calculation Configurations}
\label{sec:phonon_calc_config}
